---
title: "Compute CoP-Y"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CoPY_ComputeR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(BalanceMate)
```

## Description

The most basic indicator of body displacements is the center of pressure. It corresponds to the point on a surface where the total pressure of a person's body weight is applied. As such, it reflects the body's balance and stability during standing or movement.

Center of pressure can be computed on two axes: 

- Antero-posterior axis, a.k.a., CoP-Y
- Medio-lateral axis, a.k.a., CoP-X.

Quite counterintuitively, the CoP-Y is computed from the moments on the X-axis, and the CoP-X is computed from the moments on the Y-axis. You should imagine Moments on the X-axis as the rotation of a bar orientated following the X axis. Rotations of this bar, as you can imagine, will thus indicate leaning forward vs. backward.

Similarly, Moments on the Y-axis are rotations following the antero-posterior axis, thus indicating leaning on the right or the left. By convention, we reverse this rotation in the computation of CoP-X (i.e., we use -My, and not My). This result in a movement toward the right being a negative CoP and movement toward the left being a positive CoP. In simpler terms, if you were looking at a participant standing on a platform, this participant leaning toward your right would indicate a positive CoP. In a way, conventionnally, CoP computations are participant-centered, but observer-centered (if it makes sense).

## Example usage

The `CoPY_ComputeR` automatically compute CoP-Y from moments and forces. It is only useful if you have a full description of your moments and forces. In contrast to other functions from the package, the `CoPY_ComputeR` only requires two vectors (Moments and Forces) to output a third vector (CoP).

You collected 6 participants viewing passively some pictures for 100s.

*Note: the files are stored as Rdata to reduce the size of the package -- however the command only accepts txt files, hence it is required to first convert Rdata to text files before running the examples*

```{r}
# 1. Locate the installed extdata/ folder (contains only .RData now)
path_extdata <- system.file("extdata", package = "BalanceMate")

# 2. Find all .RData (case‐insensitive)
rdata_files <- list.files(
  path_extdata,
  pattern    = "\\.[Rr][Dd]ata$",
  full.names = TRUE
)

# 3. Convert each .RData to a .txt in a temp directory
tmpdir <- tempdir()
for (f in rdata_files) {
  load(f)  # e.g. loads object “Postural_DataA” into your session
  nm <- tools::file_path_sans_ext(basename(f))
  write.table(
    get(nm),
    file      = file.path(tmpdir, paste0(nm, ".txt")),
    sep       = ",",
    row.names = FALSE,
    col.names = FALSE,
    quote     = FALSE
  )
}


# 4. Merge from the tempdir
Data <- Merge_PosData(
  directory_path = tmpdir,              
  SampleRate     = 100,
  SessionDuration= 100
)



```

You can compute CoP-Y from this dataset using the `CoPY_ComputeR` command:

```{r}
CoPY_ComputeR(Data$Mx, Data$Fz)[1:20] # Only displaying first 20 values
```

***Note***: the `Merge_PosData` automatically computes CoP-X and CoP-Y, so if you used this command, the `CoPX_ComputeR` and `CoPY_ComputeR` are redundant.


## Further reading

Quijoux, F., Nicolaï, A., Chairi, I., Bargiotas, I., Ricard, D., Yelnik, A., Oudre, L., Bertin-Hugault, F., Vidal, P. P., Vayatis, N., Buffat, S., & Audiffren, J. (2021). A review of center of pressure (COP) variables to quantify standing balance in elderly people: Algorithms and open-access code. *Physiological reports, 9*(22), e15067. https://doi.org/10.14814/phy2.15067 

