---
title: "Compute Standard deviations of CoP displacements"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SD_CoP_ComputeR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(BalanceMate)
```


## Introduction

From a broad perspective, two types of postural indicators can be computed from the `BalanceMate` package: 

- Indicators that measure average displacements (i.e., CoP-X and CoP-Y)
- Indicators that measure variability in displacements (e.g., Sway Path Length)

Whereas the former family is useful for researchers interested in motivation (approach, and avoidance), the latter is of interest for researchers assessing loss of stability through substance use, or freezing behaviours when facing threat (among other examples).

The `SD_CoP_ComputeR` command will compute standard deviations of CoP-X and SD CoP-Y at a pre-specified time bin level (or at a participant level) from a list of CoP.

## Example Usage

You collected the data of 6 participants who viewed passively pictures for 100 seconds with a sample rate if 100Hz. You are interested in the standard deviations of CoP-Y and CoP-X displacements during the first half of the session (pleasant pictures) and the last half of the sessions (unpleasant pictures).

```{r}
path_to_data <- system.file("extdata", package = "BalanceMate")
Data <- Merge_PosData(path_to_data, SampleRate = 100, SessionDuration = 100)


BalanceMate::SD_CoP_ComputeR(data = Data, CoPX_col = "CoP_X", CoPY_col = "CoP_Y", ID = "file_name", time_col = "Time", epoch_length = 50)
```

***Note***: the `Merge_PosData` automatically computes CoP for each data points, as well as the Time course for each session.

## Further readings

See ["Processing" chapter in the BalanceMate Guide book](https://jeanmoneger.com/book/processing)


