---
title: "Visualise Stabilogram and Confidence Ellipse"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SpaghettEllipse}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(BalanceMate)
```

## Introduction

During your exploration of postural data, you might be interested in visualising CoP displacements. This can be done using stabilograms that will display CoP displacements on the medio-lateral and antero-posterior axes.

Sometimes these plots are called "Spaghetti plots", because they do look like spaghetti. However, it is important to note that spaghetti plots are also the names of other types of plots that - arguably - look slightly less like spaghetti.

The `SpaghettEllipse` command will output a spaghetti plot coupled with the X% confidence ellipse area.

## Example Usage

You collected the data of 6 participants who viewed passively pictures for 100 seconds with a sample rate if 100Hz. You now want to assess displacements for each participants. 

```{r}
path_to_data <- system.file("extdata", package = "BalanceMate")
Data <- Merge_PosData(path_to_data, SampleRate = 100, SessionDuration = 100)

SpaghettEllipse(Data, participant_id_col = "file_name", participant_id = "Postural_DataA.txt", Title = "Participant A")

SpaghettEllipse(Data, participant_id_col = "file_name", participant_id = "Postural_DataB.txt", Title = "Participant B")

SpaghettEllipse(Data, participant_id_col = "file_name", participant_id = "Postural_DataC.txt", Title = "Participant C")

SpaghettEllipse(Data, participant_id_col = "file_name", participant_id = "Postural_DataD.txt", Title = "Participant D")

SpaghettEllipse(Data, participant_id_col = "file_name", participant_id = "Postural_DataE.txt", Title = "Participant E")

SpaghettEllipse(Data, participant_id_col = "file_name", participant_id = "Postural_DataF.txt", Title = "Participant F")
```

***Note***: you can adjust the size of the Ellipse by using the `conf_level` argument.

However, you might also be interested in a specific sample of those displacements, say between the 10th second and the 90th second. This can be done using the arguments `time_col` to specify the column containing the time stamp for each data points, and the the `time_start` and `time_end` arguments to indicate the limits between which you want to plot your spaghetti.

```{r}
path_to_data <- system.file("extdata", package = "BalanceMate")
Data <- Merge_PosData(path_to_data, SampleRate = 100, SessionDuration = 100)

SpaghettEllipse(Data, participant_id_col = "file_name", participant_id = "Postural_DataA.txt", time_col = "Time", time_start = 10, time_end = 90, Title = "Participant A")

SpaghettEllipse(Data, participant_id_col = "file_name", participant_id = "Postural_DataB.txt",  time_col = "Time", time_start = 10, time_end = 90,Title = "Participant B")

SpaghettEllipse(Data, participant_id_col = "file_name", participant_id = "Postural_DataC.txt",  time_col = "Time", time_start = 10, time_end = 90,Title = "Participant C")

SpaghettEllipse(Data, participant_id_col = "file_name", participant_id = "Postural_DataD.txt",  time_col = "Time", time_start = 10, time_end = 90,Title = "Participant D")

SpaghettEllipse(Data, participant_id_col = "file_name", participant_id = "Postural_DataE.txt",  time_col = "Time", time_start = 10, time_end = 90,Title = "Participant E")

SpaghettEllipse(Data, participant_id_col = "file_name", participant_id = "Postural_DataF.txt",  time_col = "Time", time_start = 10, time_end = 90,Title = "Participant F")
```

But you might also prefer to plot more complex sway paths. For instance, if your session contains unpleasant trials and pleasant trials, you might want to compute *average sway paths across all participants across all unpleasant trials* to compare it to *average sway paths across all participants across all pleasant trials*. This can be done with a little bit of effort and data wrangling as can be seen in [the dedicated subsection of the Balance mate guide book](https://jeanmoneger.com/book/processing#visualise).


## Further readings
See the the  subsection on [Visualisation in the Balance mate guide book](https://jeanmoneger.com/book/processing#visualise).
