---
title: "Compute CoP X"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CoPX_ComputeR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(BalanceMate)
```

## Description

The most basic indicator of body displacements is the center of pressure. It corresponds to the point on a surface where the total pressure of a person's body weight is applied. As such, it reflects the body's balance and stability during standing or movement.

Center of pressure can be computed on two axes: 

- Antero-posterior axis, a.k.a., CoP-Y
- Medio-lateral axis, a.k.a., CoP-X.

Quite counterintuitively, the CoP-Y is computed from the moments on the X-axis, and the CoP-X is computed from the moments on the Y-axis. You should imagine Moments on the X-axis as the rotation of a bar orientated following the X axis. Rotations of this bar, as you can imagine, will thus indicate leaning forward vs. backward.

Similarly, Moments on the Y-axis are rotations following the antero-posterior axis, thus indicating leaning on the right or the left. By convention, we reverse this rotation in the computation of CoP-X (i.e., we use -My, and not My). This result in a movement toward the right being a negative CoP and movement toward the left being a positive CoP. In simpler terms, if you were looking at a participant standing on a platform, this participant leaning toward your right would indicate a positive CoP. In a way, conventionnally, CoP computations are participant-centered, but observer-centered (if it makes sense).

## Example usage

The `CoPX_ComputeR` automatically compute CoP-X from moments and forces. It is only useful if you have a full description of your moments and forces. In contrast to other functions from the package, the `CoPX_ComputeR` only requires two vectors (Moments and Forces) to output a third vector (CoP).

You collected 6 participants viewing passively some pictures for 100s.

*Note: the files are stored as RData to reduce the size of the package -- however the command only accepts txt files, hence it is required to first convert RData to text files before running the examples*

```{r}
# Data<-Merge_PosData("~/Desktop/BalanceMate/inst/extdata/", SampleRate = 100, SessionDuration = 100)

# If required: convert RData to text files.

# Define the files and corresponding output names
files <- list(
    "Postural_DataA" = "~/Desktop/BalanceMate/data/Postural_DataA.RData",
    "Postural_DataB" = "~/Desktop/BalanceMate/data/Postural_DataB.RData",
    "Postural_DataC" = "~/Desktop/BalanceMate/data/Postural_DataC.RData",
    "Postural_DataD" = "~/Desktop/BalanceMate/data/Postural_DataD.RData",
    "Postural_DataE" = "~/Desktop/BalanceMate/data/Postural_DataE.RData",
    "Postural_DataF" = "~/Desktop/BalanceMate/data/Postural_DataF.RData"
)

# Loop through each file, add a blank row with spaces, and save as .txt
for (name in names(files)) {
    # Load the .RData file
    load(files[[name]])
    
    # Dynamically get the object loaded (assuming it's named the same as the file)
    data <- get(name)
    
    # Add a "blank" row (spaces for all cells) at the top
    #blank_row <- rep(" ", 6)  # Create a row of spaces with the same number of columns
    #data <- rbind(blank_row, data)
    
    # Write to a .txt file with the same name
    write.table(data, file = paste0(name, ".txt"), sep = ",", row.names = FALSE, col.names = FALSE, quote = FALSE)
}

Data<-Merge_PosData(".", SampleRate = 100, SessionDuration = 100)


```

You can compute CoP-X from this dataset using the `CoPX_ComputeR` command:

```{r}
CoPX_ComputeR(Data$My, Data$Fz)[1:20] # Only displaying first 20 values
```

***Note***: the `Merge_PosData` automatically computes CoP-X and CoP-Y, so if you used this command, the `CoPX_ComputeR` and `CoPY_ComputeR` are redundant.


## Further reading

Quijoux, F., Nicolaï, A., Chairi, I., Bargiotas, I., Ricard, D., Yelnik, A., Oudre, L., Bertin-Hugault, F., Vidal, P. P., Vayatis, N., Buffat, S., & Audiffren, J. (2021). A review of center of pressure (COP) variables to quantify standing balance in elderly people: Algorithms and open-access code. *Physiological reports, 9*(22), e15067. doi: 10.14814/phy2.15067 


